---
title: "Strawberries Assignment"
author: "Jin Wen Lin"
format: pdf
editor: visual
---

```{r}
#| label: load libraries and set options
#| warning: false
#| message: false
#| 

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
library(tidyverse)
```

First, read the data and see the overall summary.

```{r}
#| label: read data - glimpse 

strawberry <- read_csv("strawberries25_v3.csv")

glimpse(strawberry)
```

```{r}
#|label: function def - drop 1-item columns (code from lecture)

drop_one_value_col <- function(df){   ## takes whole dataframe
drop <- NULL  

## test each column for a single value
for(i in 1:dim(df)[2]){     
if((df |> distinct(df[,i]) |> count()) == 1){
drop = c(drop, i)
} }

## report the result -- names of columns dropped
## consider using the column content for labels 
## or headers 

if(is.null(drop)){return("none")}else{

   print("Columns dropped:")
   print(colnames(df)[drop])
   strawberry <- df[, -1*drop]
   }
}


## drop the columns with NAs

strawberry <- drop_one_value_col(strawberry)

```

Separate the data item column into columns of strawberries, Measure, and Bearing type.

```{r}
#| label: Census Data
# strawberry census
strawberry_cens<- strawberry %>% filter(Program == "CENSUS")
data_item <- strawberry_cens %>% distinct(`Data Item`)
data_item

## Separate the Data Item column
strawberry_cens <- strawberry_cens %>% 
  separate_wider_delim(  cols = `Data Item`,
                         delim = " - ",
                         names = c("strawberries",
                                 "Category"),
                         too_many = "error",
                         too_few = "align_start"
                       )

strawberry_cens <- strawberry_cens %>% 
  separate_wider_delim(  cols = `Category`,
                         delim = " ",
                         names = c("Measure",
                                 "Bearing_type"),
                         too_many = "merge",
                         too_few = "align_start"
                       )
# remove commas in the Measure Column
strawberry_cens <- strawberry_cens %>% 
  mutate(Measure = gsub("\\,", "", Measure))

# organic census data
organic_cens <- strawberry_cens %>% filter(str_detect(strawberries, "ORGANIC")==TRUE)

# non-organic census data
non_organic_cens <- strawberry_cens %>% filter(str_detect(strawberries, "ORGANIC")==FALSE)


```

Separate the Domain Category column into three new columns with chemical name, chemical type, and chemical code.

```{r}
#|label: Strawberry Survey Data

strawberry_survey <- strawberry %>% filter(Program == "SURVEY")
strawberry_survey %>% distinct(strawberry_survey$`Domain Category`)
# remove CHEMICAL in the Domain Category
strawberry_survey <- strawberry_survey %>% 
  mutate(`Domain Category` = gsub("CHEMICAL, ", "", `Domain Category`))
# separate them into chemical name, chemical type, and chemical code
strawberry_survey <- strawberry_survey %>% 
  separate_wider_delim(  cols = `Domain Category`,
                         delim = ": ",
                         names = c("Type",
                                 "Others"),
                         too_many = "error",
                         too_few = "align_start"
                       )

strawberry_survey <- strawberry_survey %>% 
  separate_wider_delim(  cols = Others,
                         delim = " = ",
                         names = c("Name",
                                 "Chemical Code"),
                         too_many = "error",
                         too_few = "align_start"
                       )
# remove brackets
strawberry_survey <- strawberry_survey %>% 
  mutate(Type = gsub("\\)", "", Type))

strawberry_survey <- strawberry_survey %>% 
  mutate(Type = gsub("\\(", "", Type))

strawberry_survey <- strawberry_survey %>% 
  mutate(`Chemical Code` = gsub("\\)", "", `Chemical Code`))



```

```{r}
#|label: Check the Value Column

# organic census data Value
value_1 <- organic_cens %>% distinct(Value)
value_1
# change Value into numbers
organic_cens$Value <- as.numeric(gsub(",", "", organic_cens$Value))

# non-organic census data Value
value_2 <- non_organic_cens %>% distinct(Value)
value_2
# change Value into numbers
non_organic_cens$Value <- as.numeric(gsub(",", "", non_organic_cens$Value))

# survey data
# change Value into numbers
strawberry_survey$Value <- as.numeric(gsub(",", "", strawberry_survey$Value))
```

While turning the value column into numbers or doubles instead of characters, there are some values automatically turned into NA such as notations like (D) and (Z) etc. The different notations do have different meanings behind them but now they are all in NAs.

```{r}
# write the survey data into csv
write.csv(strawberry_survey, "strawberry_survey.csv", row.names = FALSE)
```
